\documentclass[12pt]{article}

\title{Luke Palmer}
\author{Software Engineer}
\date{720-837-6741 \\ \texttt{lrpalmer@gmail.com}}

\begin{document}
\maketitle

I have twelve years of programming experience with focus on software
architecture and abstraction, programming languages\footnote{Embedding,
using, and designing.}, and automated testing and quality assurance.  I
have nine years of technical communication experience (on the Perl 6
project and as a teaching assistant). I have a fundamental desire to
design clean, elegant, and modular code.

\section*{Accomplishments}

\begin{itemize}
\item Worked with Larry Wall as a member of the Perl Cabal (Perl 6
Design Team) from 2004 to 2006.
\item Implemented many experimental Haskell modules, including a FRP
game library called Fregl and an Arrow implementing a model of Quantum
computation (\texttt{quantum-arrow} on hackage). \\
\verb|http://luqui.org/blog/archives/2008/02/15/frp-and-parallelism/| \\
\item Implemented the Perl Grammar Engine in Perl 6, an essential piece
-of the bootstrapping process, using a very modular design (``virtual
classes''). \\
\verb|http://svn.pugscode.org/pugs/ext/Parse-Rule/|
\item Authored 15 Perl 5 modules including a large predicate logic
engine, several object-oriented design tools, an attribute grammar
engine (complex data structure processing tool), among other things, all
with extensive documentation and test suites. \\
\verb|http://search.cpan.org/~lpalmer/|
\item Authored and co-authored many small, cross-platform games, some of
which can be found at \verb|http://luqui.org/stuff|.  My largest
contributions to the projects there were: the real-time fluid dynamics
engine for Ichor, the fast automated solver for Minesweeper Infinity (to
ensure that the user never gets an unsolvable board), and the
transactional logic engine for Telegnosis (well, all of Telegnosis,
actually).
\item Presently working on a 2D shooter on non-euclidean manifolds in my
free time, which looks really cool (I promise).
\item Presented as an ``emerging topics'' speaker at the O'Reilly Open
Source Convention 2004 a talk entitled ``A Language for Games''.  \\
\verb|http://conferences.oreillynet.com/cs/os2004/view/e_sess/4969|
\end{itemize}

\section*{Career History/Education}

\begin{itemize}
\item Software engineer at NetDevil game studio from February to
September 2007, working on the LEGO Universe project (a large scale MMO)
and the LEGO Web Racers (an small asynchronous MMO in Flash with
Actionscript 3).
\item First place in Colorado (of 12 teams) and 4th in region
(45 teams) in the 2006 ACM programming competition.
\item BA/MA student of mathematics at the University of Colorado
from 2002-2006, presently on some sort of hiatus.  GPA: 3.32; Major
(mathematics) GPA: 3.80.
\item Worked on the large, open source Perl 6 project continuously from
2001 to 2006.
\item Teaching Assistant in Physics 2 at CU in 2006 under Dr. Steven
Pollock.
\item Teaching Assistant in Introductory Computer Science at CU in 2006
under Dr. Susanne Sherba.
\item Worked on MVT (mathematics visualization toolkit) for the summer
of 2005. \verb|http://amath.colorado.edu/java/index.php|
\item Worked on a medical natural language processing project in 2004 with
Dr. Robert Bruegel and Dr. Ngo Thanh Nhan (not associated with the
University of Colorado).
\item Teaching Assistant in Calculus 2 at CU in 2004 under Dr. Mark
Ablowitz.
\item Worked on a medical expert system in 2003 with Dr. Robert Bruegel.
\item Teaching Assistant in Physics 2 at CU in 2003 under Dr. Mike
Dubson.
\end{itemize}

\section*{Qualifications}

\begin{itemize}
\item 10 years experience with C and C++; I know all weird the nooks and
crannys of C++, and can use them in order to achieve the best design.
\item 3 years of experience with Haskell; a very good knowledge of the
language and libraries.  
\item Experience with many other languages, including Perl,
Ruby, Lua, Python, Java, Lisp, Prolog, Curry, Scala, ....  I
I can learn new languages very quickly.
\item Experience with the design and implementation of general purpose
programming languages and domain specific languages; experience with
code correctness proofs.
\item Lots of experience with Subversion and some with CVS and Darcs.
\item Experience with JUnit, Test::More, Quickcheck, and building
unit testing frameworks with Test::Builder.
\item Fundamental desire for clean, abstract software design.  Lots of
experience refactoring code into such designs.
\end{itemize}

\section*{References}

\begin{itemize}
\item Larry Wall. Mountain View, CA.  650-691-9038 (home).
\verb|larry@wall.org|.
\item Ivan Woehr (my lead programmer at NetDevil);
\texttt{ivan@netdevil.com}.
\item Dr. Robert Bruegel.  Lafeyette, CO.  303-499-1685.
\verb|rbruegel@aol.com|
\end{itemize}

\end{document}
