\documentclass[12pt]{article}

\usepackage{setspace}

\begin{document}
\noindent Luke Palmer \\
2005-09-13 \\
CSCI 3702 \\
Polonsky, Blake, Braun, and Heeger: Neuronal activity in human primary
visual cortex correlates with perception during binocular rivalry.

\onehalfspace

This paper describes an experiment studying binocular rivalry.  Each
subject was first shown a different image to each eye: a different solid
color in each (with differing contrasts), together with darkened
stripes, vertical in one eye, horizontal in the other.  The subject was
asked to push one of three buttons at any time: one button if he saw
horizontal stripes, another if he saw vertical stripes, and a third if
the image was ``composite''; that is, if less that 75\% of a single
image could be seen.  A portion of the sequence was then played back to
the subject in a non-rivalrous way (conflicting colors to each eye, but
a striped pattern would only be present in one eye at a time), and the
subject was asked to press the same buttons.

The experiment measured activity in V1 during these experiments with
fMRI, and found that V1 activity was coupled with the reported
perceptual alternations.  However, the fluctuations in nearby visual
areas was roughly the same.  The experimentors conclusively stated that
these results ``could mean that neuronal events underlying rivalry are
initiated in V1 and then propagated to later areas, or that those
neuronal events are initiated at later stages of processing and then
propagated via feedback to V1.  It is also possible that both processes
occur.''

This passage made me wonder:

\begin{quote}
\singlespace
To ensure that our conclusions did not depend critically on the
assumptions of the model, we did a complementary analysis ....  This
analysis produced similar state amplitudes to the model fits.
\end{quote}

It seems like they fit the model to the data, and then made sure that
the model in fact fit the data.  This seems silly and doesn't ensure
much of anything.  Maybe I don't understand the process.

\end{document}
