\documentclass[12pt]{article}

\usepackage{fancyhdr}
\usepackage{amsmath}

\lhead[\fancyplain]{Luke Palmer}
\chead[\fancyplain]{2006-10-05}
\rhead[\fancyplain]{LING 3430}
\pagestyle{fancyplain}

\begin{document}
\section*{5.1}
The following are stative: \textit{comprise own possess know resemble
lack\footnote{Perhaps: ``He is lacking a sense of accomplishment'' seems
to be a counterexample.} prefer}

\section*{5.2}
\begin{itemize}
\item (D) I am really admiring nature today.
\item (S) I admire Keanu Reeves for his excellent presentation of sarcasm.
\item (D/S) \textit{equal?}
\item (D) A large bear appeared.
\item (S) The large bear appeared angry.
\item (D) Hold your horses!
\item (S) \textit{hold?}
\item (D) The firefighters were containing the fire.
\item (S) ``That bottle contains poison.''
\item (D) I reached for the bottle anyway.
\item (S) \textit{reach?}
\item (D) The attack of the clones cost many lives.
\item (S) ``The Attack of The Clones'' costs twenty dollars.
\item (D) John smelled the pigs.
\item (S) John smelled like a pig.
\end{itemize}

\section*{5.3}
The following are punctual verbs: \textit{ring\footnote{As in ``the
phone rang all night'', but not as in ``my ears rung all night''.} tap
sigh twitch sob}

\section*{5.4}
\begin{description}
\item[Dismantle a car] is telic.  \textbf{Dismantling cars} is atelic.
\item[Swim] is atelic.  \textbf{Swim a lap} is telic.
\item[Rig an election] is telic.  \textbf{Rig elections} is atelic (this
is easy :-).
\item[Put out fires] is atelic.  \textbf{Put out a fire} is telic.
\item[Ripen] is telic.  I cannot think of a way to make this atelic.
\item[Direct movies] is atelic.  \textbf{Direct a movie} is telic.
\item[Sink a beer] is telic (what does that mean?).  I'm guessing
  \textbf{sink beers} is atelic, but I don't know this sense of ``sink.''
\end{description}

\section*{5.8}
\begin{description}
\item[a.]  Epistemic: ``it is possible that beer is not served to
minors''; Deontic: ``you are not allowed to serve beer to minors''.
\item[b.]  Epistemic: ``you are capable of going home now''; Deontic:
``I will allow you to go home now''.
\item[c.]  Epistemic: ``we are capable of taking the exam early'';
Deontic: (not sure) ``we are allowed to take the exam early''.
\item[d.]  Epistemic: ``at no point in the future will you have left
this island''; Deontic: ``I will never allow you to leave this island''.
\item[e.]  Epistemic: ``it is probable that we will get to the hotel by
nine''; Deontic: ``if we don't get to the hotel by nine, someone will be
angry''.
\end{description}

\section*{5.10}
\begin{description}
\item[a.] Induction: ``evidently''.
\item[b.] Deduction: ``should''.
\item[c.] Sensory evidence: ``look like''.
\item[d.] Hearsay evidence: ``Apparently''.
\item[e.] Sensory evidence: ``sounds''.
\item[f.] Induction: ``must''.
\item[g.] Hearsay evidence: ``allegedly''.
\item[h.] Belief: ``suppose''.
\end{description}
\end{document}
